---
output: github_document
bibliography: src/references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# BFun

<!-- badges: start -->
[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active) [![License: GNU-3](https://img.shields.io/badge/license-GNU--3-blue.svg)](https://cran.r-project.org/web/licenses/GNU-3) [![](https://img.shields.io/github/last-commit/JSartini/BFun.svg)](https://github.com/JSartini/BFun/commits/main) [![](https://img.shields.io/badge/devel%20version-0.1.0-blue.svg)](https://github.com/JSartini/BFun)
<!-- badges: end -->

The goal of BFun is to provide tools necessary to fit Bayesian Functional (Mixed) Models where random effects are represented using stochastic eigenfunctions. This type of joint modeling accounts for all potential sources of uncertainty, providing valid inferences even when signal to noise ratio is relatively small. The package is oriented around a Markov Chain Monte Carlo approach, but provides an experimental variational implementation as well (still undergoing testing).

For the statistical insights behind these models, refer to [@Sartini08122025] and [@sartini_bayesian_2025].

## Installation

You can install BFun from GitHub using the `remotes` package as follows:

``` r
install.packages("remotes")
library(remotes)
remotes::install_github("JSartini/BFun")
```

## Example

This is a basic example showing how to fit a Bayesian functional model using BFun on the DTI data from the `refund` package. The model is a simple multilevel FPCA here, as no covariates are included in the formula.

```{r example, eval=FALSE}
library(BFun)
library(refund)
library(tidyr)

# Testing data
data(DTI)
fit_data = drop_na(DTI)

# Fit the model
Mod = bfmm(cca ~ 1, data = fit_data, 
           id = "ID", visit = "visit", 
           n_chains = 4, n_cores = 4, 
           n_iter = 1000, n_burnin = 1000, 
           alpha = 0.05)

# Raw samples for inference on any desired functional
Mod$Samples

# Estimates
Mod$Estimates

# Uncertainty bounds at given alpha level
Mod$Inferences

# Diagnostics (variance explained, MCMC convergence)
Mod$Diagnostics
```

# References
